---
title: "Graphing with ggplot"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r}
library(ggplot2)
```

####Read data
```{r}
foodData=read.csv('Data/food-texture.csv',
header=TRUE, row.names = 1)
print(foodData)
str(foodData)
```

Data types in R:character,integer,numeric,boolean,complex
Data structures:vector(1-D  data structure with all with same data type-homogeneous)
                list(1-D -heterogeneous)-use list function
                matrix(2-D -homogeneous)
                dataframe(2-D heterogeneous)
                factor,array
typeof-gives raw datatype
class-gives datatype of data stored
str-gives structure of an object
index starts from 1 in R language


####Attributes of the dataframe
```{r}
nrow(foodData)
ncol(foodData)
rownames(foodData)
colnames(foodData)
```
####Accessing elements of a dataframe
```{r}
foodData$Oil #prints values of oil data-vector-all theese first three
foodData[['Oil']]
foodData[,'Oil']
str(foodData['Oil']) #dataframe
```

####Factor data structure in R 
```{r}
gender=c('Male','Female','Female','Male','Female','Male','Male')
print(gender)
gender_categorical=factor(gender)
print(gender_categorical)
str(gender_categorical)
levels(gender_categorical)
```

####Add a new column called Taste with levels (1)poor (2)fair (3)good (4)excellent
```{r}
#foodData['Taste']=sample(c('poor','fair','good','excellent'), nrow(foodData), replace=TRUE,p=c(0.25,0.25,0.25,0.25))
#head(foodData)
```

####Write the updated dataframe to food-texture
```{r}
#write.csv(foodData, 'Data/food-texture.csv')
```

####Create lists of continuous and categorical feature names
```{r}
continuous_cols=c('Oil','Density','Fracture','Hardness')
categorical_cols=c('Crispy','Taste')
```

####Convert all categorical columns (that we know) into factor type in one short using lapply-returns list with column by column values
```{r}
foodData[categorical_cols]=
lapply(foodData[categorical_cols],factor)
str(foodData)
```

####Create a ggplot canvas
```{r}
p=ggplot(data=foodData)
p
```

####Visualize and understand the relationship between two continuous variables(scatter plot): Oil and Density
```{r}
p=ggplot(data=foodData)+
  geom_point(aes(x=Density,y=Oil,color=Crispy))
p
```

```{r}
p=ggplot(data=foodData)+
  geom_point(aes(x=Density,y=Oil,color=Crispy))
p
```

####Visualize and understand the relationship between a continuous and a categorical feature(box plot): Density and Crispy
```{r}
p=ggplot(data=foodData)+
  geom_boxplot(aes(x=Crispy, y=Density))
p
```

####Understand what is inside a box plot
```{r}
#foodData$Density
#foodData$Crispy==10
#foodData[foodData$Crispy==10, ]
#foodData[foodData$Crispy==10 & foodData$Oil<=17,]
#foodData[foodData$Crispy==10, c('Density','Hardness')]
density10=foodData[foodData$Crispy==10,'Density']
density11=foodData[foodData$Crispy==11,'Density']
str(density10)
mean(density10)
density10-mean(density10) #Broadcasting-mean centered vector
(density10-mean(density10))^2 #Squared mean centered vector
mean((density10-mean(density10))^2) #mean of the squared mean-centered vector-varience
sqrt(mean((density10-mean(density10))^2))#Standard deviation of the vector

var(density10) #varience
sd(density10) #standard deviation
```


```{r}
# Load the mpg dataset
data('mpg')
mpgData = mpg

# Print the first five rows (or samples) in the data frame
head(mpgData, 5)
```
```{r}
# Initiate the ggplot() function binding to the car data frame
ggplot(data = mpgData)

# Create a plot object
p1 = ggplot(data = mpgData)

# Use the aes() function to specify the aesthetic mapping, that is, which variables
# should be plotted
p1 = ggplot(data = mpgData, aes(x = disp, y = hwy))

# Use the geom_ type functions to add geometric elements
p1 = ggplot(data = mpgData, aes(x = displ, y = hwy)) +
  geom_point()

# Add labels and title
p1 = p1 + labs(x = 'Displacement (litres)', y = 'Mileage (MPG)', title = 'Mileage vs. Displacement')
p1
```


```{r}
# Map aesthetics to variables

# Map the color aesthetic to the class variable
p2 = ggplot(data = mpgData) +
  geom_point(mapping = aes(x = displ, y = hwy))
p2
```

```{r}
# Map aesthetics to variables

# Map the size (or alpha or shape) aesthetic to the class variable
p3 = ggplot(data = mpgData) +
  geom_point(aes(x = displ, y = hwy))
p3
```

```{r}
# Set aesthetic manually

# Mark the points blue in color, square in shape, and fixed size
p4 = ggplot(data = mpgData) +
  geom_point(aes(x = displ, y = hwy))
p4
```

```{r}
# Add additional variables using facets

# Plot mileage w.r.t. each class individually
ggplot(data = mpgData) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(, nrow = 2)
```

```{r}
# Plot mileage w.r.t drive train and number of cylinders
ggplot(data = mpgData) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid()
```

```{r}
# A quick exercise on facet plotting with filtering of samples
# Investigate the number of levels in the cyl feature
factor()

# How many cars of each cyl type are there?
mpgData %>% 

# Filter samples with 3 and 5 cylinder cars (which are very rare)
# mpgData = mpgData %>% filter(!(? %in% c(?, ?)))
# 
# head(mpgData, 5)
# 
# # Map the color aesthetic to the cyl variable
# p5 = ggplot(data = mpgData) +
#   geom_point(mapping = aes(x = displ, y = hwy))
# p5
```


